Limitations on the Two-Infall Scenario in the Era of Large Stellar Age Catalogs

### 2025-04-24 src/tex/ms.tex
Updates from Overleaf

% \todo{Hook, summary of alpha-bimodality.}
\lettrine{G}{alactic chemical evolution} (GCE) studies aim to explain the observed stellar abundance patterns in the Milky Way by modeling the star formation history and evolution of the Galaxy. A long-standing paradigm of GCE is that the metallicity of the interstellar medium (ISM) increases over time due to supernova enrichment from successive generations of stars. In this view, one feature of the Milky Way disk that is difficult to explain is the so-called ``$\alpha$-bimodality'': the presence of two populations of stars at similar metallicity but separated by their \aFe ratio \citep[e.g.,][]{bensby_exploring_2014}. The high-$\alpha$ sequence consists of old stars \citep[$\gtrsim9$ Gyr;][]{pinsonneault_apokasc-3_2025} with super-Solar \aFe and is associated with the kinematic thick disk \citep[e.g.,][]{fuhrmann_nearby_1998}, while the low-$\alpha$ sequence is younger, with approximately Solar \aFe, and is associated with the thin disk. The $\alpha$-bimodality is present across the Galactic disk, but the relative strength of the high- and low-$\alpha$ sequences varies by location \citep{hayden_chemical_2015}.

An explanation for the $\alpha$-bimodality has yet to be broadly accepted in the GCE community \todo{(reword)}. An $\alpha$-bimodality is not a universal feature in simulated Milky Way-mass galaxies \todo{(citations; also mention external galaxies?)}, so its presence and characteristics in our own galaxy may provide clues to its unique evolutionary history. GCE models that attempt to solve this problem generally fall into two camps. Some explain the $\alpha$-bimodality as a result of secular processes, such as the radial migration of stars and the inside-out growth of the disk \todo{(citations)}. Others argue that the two sequences point to a bursty star formation history, perhaps driven by multiple discrete accretion events \citep[e.g.,][]{chiappini_chemical_1997,mackereth_origin_2018,spitoni_beyond_2023} or a change in the star formation efficiency \citep[e.g.][]{conroy_birth_2022}.

% Summary of two-infall literature
The two-infall model of chemical evolution was proposed by \citet{chiappini_chemical_1997} to explain the origin of the high- and low-$\alpha$ disks. Though the model has been revised and refined since, the basic premise remains the same: the rate of gas infall onto the Galaxy is described by two consecutive, exponentially declining bursts. The relatively low infall rate between the two bursts leads to a hiatus in the star formation rate, allowing the gas abundance to evolve between the high- and low-$\alpha$ sequences while producing few stars in between. 
% By varying the timescale of the second infall, the model can produce inside-out disk growth and a radial abundance gradient.
The initial model of \citet{chiappini_chemical_1997} successfully reproduced the available abundance data at the time, which were largely confined to the Solar neighborhood.

@@

% \todo{Observational evidence other than $\alpha$-bimodality.} 
The chemical and kinematic separation of the high- and low-$\alpha$ disks remains the primary observational evidence behind the two-infall model. \citet{spitoni_remind_2024} argued that the observed gap between the sequences in \aFe, contrasted with their overlap in $[\alpha/{\rm H}]$, indicates a haitus in the star formation history, which is a natural consequence of the two-infall model. In different samples, \citet{nissen_high-precision_2020} and \citet{nataf_accurate_2024} observed multiple sequences in the local age--metallicity relation, which would naturally be explained by the two-infall scenario. Many two-infall studies have also reproduced the metallicity gradient, the local surface densities of stars and gas, and the local star formation and supernova rates \todo{(citations)}, although these observables can also be matched by other models \todo{(citations)}.
% \citet{nissen_high-precision_2020} measured abundances and estimated ages for nearby Solar-type stars and reported two sequences in the age--metallicity distribution which could be explained by the two-infall model \todo{(rewrite)}. \citet{nataf_accurate_2024} also observed multiple sequences in the age--metallicity relation in their sample of $\sim400,000$ local subgiants, although the break between them is not as clean as in \citet{nissen_high-precision_2020}.

% \todo{Outflows.} 
Most previous studies of the two-infall model do not include mass-loaded outflows. Some hydrodynamic simulations of Galactic fountains ejected by CC SNe have shown that ejected material falls back onto the disk on relatively short timescales \citep{spitoni_galactic_2008,spitoni_effects_2009} and close to their point of origin \citep{melioli_hydrodynamical_2008,melioli_hydrodynamical_2009}, suggesting the effect on GCE could be minimal. However, other simulations of MW-like galaxies with different feedback prescriptions do produce mass-loaded outflows \todo{(citations)}. Empirically, mass-loaded outflows have been observed in nearby starburst galaxies \todo{(citations)} and dwarf galaxies \todo{(citations)} but not MW-like systems due to current detection limits \todo{(citations)}. We adopt an {\it ad hoc} prescription for outflows as a function of radius in order to balance different yield sets and reproduce the radial metallicity gradient. \todo{Trim down, move some to methods.}

@@

% \todo{Summary of secular evolution models which claim alpha-bimodality.} 
In contrast to the two-infall scenario, a number of studies have attempted to explain the $\alpha$-bimodality through purely secular processes. 
By incorporating stellar ``churning'' (changes in guiding-center radius) and ``blurring'' (changes in orbital eccentricity) as well as radial gas flows in a multi-ring model, \citet{schonrich_chemical_2009} successfully produced distinct high- and low-$\alpha$ sequences through secular processes alone; however, they failed to reproduce the overlap in metallicity between the two that would be a key finding of later surveys \citep[e.g.,][]{bensby_exploring_2014}. \citet{kubryk_evolution_2015} produce two parallel sequences in \aFe vs [Fe/H] by making a simple age cut to the results of their multi-zone GCE model, though because they do not present stellar \aFe distributions it is hard to determine whether their distributions are truly bimodal. In the model of \citet{sharma_chemical_2021}, a sharp transition in stellar \aFe for $8-12\Gyr$ old stars plus radial migration are the key ingredients for predicting a MW-like bimodality; however, their approach to the chemical evolution is not {\it ab initio}, meaning that instead of integrating the enrichment equations for their adopted star formation history, they simply assume a functional form for the abundance evolution over time \todo{(rewrite)}. \todo{Sharma basically says that it's possible to have a continuous star formation history and reproduce the $\alpha$-bimodality.} \citet{chen_chemical_2023} confirmed the findings of \citet{sharma_chemical_2021} in an {\it ab initio} multi-zone GCE model (with a two-phase ISM and radial gas flows but no outflows) and were able to produce a bimodal \aFe distribution; however, some discrepancies remain, including that the metallicity of the high-$\alpha$ sequence is too low and not extended enough. \citet{chen_chemical_2023} found that the abundance gap is caused by the simultaneous decline in SFR and \aFe. 
\todo{\citet{prantzos_origin_2023}.}
\citet{chen_recent_2025} argue that the double age--metallicity sequences observed by \citet{nissen_high-precision_2020} can be explained by a smooth star formation history with inside-out growth: the older sequence consists of stars which have migrated from the inner galaxy, where the star formation rate peaked early, while the younger sequence was formed locally. \todo{Trim this WAY down.}

% \todo{Introduce equilibrium model.} 
As stellar age estimation techniques have improved over recent years, large catalogs have become available with ages for hundreds of thousands or even millions of stars from a wide swath of the Galaxy. In a challenge to the traditional view of GCE, which expects the ISM metallicity to continually increase over time, \citet{johnson_milky_2024} examined the age--metallicity relation at different radii from the astroNN catalog \citep{mackereth_dynamical_2019} and found that the mode of the metallicity distribution at a given radius is nearly independent from age over the past $\sim9\Gyr$. They propose an ``equilibrium scenario'' in which the local metallicity is driven by the ratio of star formation to accretion at a given radius, which evolves to a constant over $\sim\Gyr$ timescales. Whether the equilibrium metallicity is driven by outflows, as proposed by \citet{johnson_milky_2024}, or by other factors such as a radial gas flow, the current data suggest that the gas abundance across the Galaxy has evolved very little over most of the thin disk lifetime.

Recently, \citet{dubay_galactic_2024} studied the effect of the SN Ia delay-time distribution in the context of several different star formation histories, including a two-infall model. While the results did not greatly depend on the choice of SFH, the authors \todo{(we?)} noted that the two-infall model, as implemented, produced a stellar [O/Fe] distribution with three distinct peaks, at variance with the data.

% \todo{Goal and summary of paper.} 
We run multi-zone GCE models with a two-infall accretion history, radially-dependent mass-loaded outflows, and a prescription for radial migration tuned to a hydrodynamical simulation. We investigate the impact of the scale of SN yields and outflows, the strength of radial migration, the enrichment of the circumgalactic medium, and the local disk mass surface density ratio on the GCE model predictions. We compare our results to abundance distributions across the disk from APOGEE DR17, and to age--abundance relations from multiple age catalogs. We describe our observational sample in Section \ref{sec:observational-sample}, and we detail our chemical evolution models and parameter selection in Section \ref{sec:methods}. We compare our multi-zone model predictions to the data in Section \ref{sec:multizone-results}. We discuss in Section \ref{sec:discussion} and conclude in Section \ref{sec:conclusions}.

@@

We use an outflow mass-loading factor because it allows us to test a range of yield scales while maintaining our desired endpoint of $Z_{\rm final}\approx Z_\odot$ \todo{(rewrite)}.

@@

The infall rate is described by two successive, exponentially declining bursts in time. The first infall component induces the formation of the thick disk, and the second component produces the thin disk. At a given galactocentric radius $R_{\rm gal}$, the un-normalized form of the infall rate is
\begin{equation}
    \label{eq:twoinfall-ifr}
    f_{\rm in}(t|R_{\rm gal}) = e^{-t/\tau_1} + f_{2/1}(R_{\rm gal}) e^{-(t-t_{\rm max})/\tau_2},
\end{equation}
where $\tau_1$ and $\tau_2$ are the first and second infall timescales, respectively, $t_{\rm max}$ is the onset of the second infall and thus the time of maximum gas infall, and $f_{2/1}$ is the ratio of the second infall amplitude to the first. We calculate $f_{2/1}$ for each zone such that the resulting stellar density profile follows a two-component disk, with the surface density ratio of the thick and thin disks given by
\begin{equation}
    f_\Sigma(R) \equiv \frac{\Sigma_1(R)}{\Sigma_2(R)} = f_{\Sigma,0} e^{R(1/R_2 - 1/R_1)}.
\end{equation}
We adopt a thick disk scale radius of $R_1=2.0\kpc$, a thin disk scale radius of $R_2=2.5\kpc$, and a central surface density ratio of $f_{\Sigma,0}=0.27$ \citep{bland-hawthorn_galaxy_2016}. In the Solar neighborhood, this produces a thick-to-thin disk mass ratio of $f_\Sigma(R_\odot)=0.12$. By comparison, \citet{mackereth_age-metallicity_2017} finds $f_\Sigma(R_\odot)=0.18\pm0.05$ from APOGEE. \citet{fuhrmann_local_2017} derive $f_\Sigma(R_\odot)=0.57$ from a very small volume (25 pc radius). \todo{Compare with other two-infall papers, which generally have a higher thick-to-thin ratio.} \citet{spitoni_galactic_2020} explores $f_\Sigma(R_\odot)\approx0.23-0.31$, \citet{palla_chemical_2020} assumes $f_\Sigma(R_\odot)=0.23$, \citet{spitoni_apogee_2021} derives $f_\Sigma(R_\odot)=0.18$ (again from APOGEE), \citet{spitoni_remind_2024} assumes $f_\Sigma(R_\odot)=0.4$.

@@

In our multi-zone models, we vary the second infall timescale with radius to produce inside-out growth of the disk. Previous multi-zone two-infall studies \citep[e.g.,][]{chiappini_abundance_2001,palla_chemical_2020} scale $\tau_2$ linearly with radius, with $\tau_2\approx1\,{\rm Gyr}$ in the inner disk and $\tau_2=7\,{\rm Gyr}$ at the Solar annulus. This prescription was adopted to match the metallicity distribution of the Solar neighborhood and the bulge in the absence of mass-loaded outflows \citep{romano_mass_2000}. We instead adopt an exponential $\tau_2 - R_{\rm gal}$ relation, with $\tau_2(R_\odot)=15\,{\rm Gyr}$ at the Solar annulus and a scale radius $R_{\tau_2}=7\kpc$. This is similar to the star formation history timescale of \citet{johnson_stellar_2021}, which was based on the  stellar age gradients in Milky Way-like spirals observed by \citet{sanchez_spatially_2020}. We also run models with a linear prescription and with a uniform value for $\tau_2$ and find little difference in our key results.

Finally, panel (c) of Figure \ref{fig:twoinfall-parameters} shows that the time of maximum infall $t_{\rm max}$ (c) strongly affects the overall stellar abundance distribution for values $t_{\rm max}\leq2\,{\rm Gyr}$, but in this case the gas tracks do not produce the characteristic abundance loop. For $t_{\rm max}>2\,{\rm Gyr}$, varying $t_{\rm max}$ results in a minor shift to the mean of the MDF and little change to the [O/Fe] distributions, even though the abundance tracks in [O/Fe]--[Fe/H] space appear very different. The value of $t_{\rm max}$ also slightly adjusts the ISM abundance endpoint, as a longer $t_{\rm max}$ means the chemical evolution ``reset'' from the second infall occurs closer to the present day (see discussion in Section \ref{sec:age-abundance}. We adopt a fiducial value of $t_{\rm max}=4.2\,{\rm Gyr}$ so that the second infall occurs at a lookback time of 9 Gyr, younger than the median age of the thick disk in the APOKASC-3 catalog of $9.14\pm0.05\,{\rm Gyr}$ \citep{pinsonneault_apokasc-3_2025}. This value is generally in line with previous two-infall studies \citep[e.g.,][]{nissen_high-precision_2020,spitoni_galactic_2020,spitoni_apogee_2021}. \todo{How does this compare to the age of GSE?}

@@

\begin{figure*}
    \centering
    \includegraphics{figures/spitoni_comparison.pdf}
    \caption{Chemical abundance tracks and distributions from one-zone models which use our fiducial parameters (solid curves) and the MCMC-derived parameters from \citet[][dashed curves]{spitoni_apogee_2021} at three different galactocentric radii. All models use the $y/Z_\odot=1$ yield set and corresponding outflow prescription (see Table \ref{tab:yields}). The time markers and marginal panels are the same as in Figure \ref{fig:twoinfall-parameters}. We also include a 2-D heatmap of the APOGEE stellar abundances in each panel for reference. APOGEE stars are restricted to (a) $2\leq R_{\rm gal}<6$ kpc, (b) $6\leq R_{\rm gal}<10$ kpc, and (c) $10\leq R_{\rm gal}<14$ kpc, and $0\leq |z|<2$ kpc in each panel.}
    \label{fig:spitoni-comparison}
\end{figure*}

\todo{Highlight differences between our models and those of \citet{spitoni_apogee_2021} in Figure \ref{fig:spitoni-comparison}.}

@@

The two-infall model generically predicts {\it three} peaks in the [O/Fe] distribution which correspond to the high-$\alpha$ sequence, the abundance ``turn-over'' after the second infall, and finally the late-time low-$\alpha$ sequence. Figure \ref{fig:ofe-distribution-yields} compares [O/Fe] distributions from across the Galactic disk produced by models with the \yZ{1} and \yZ{2} yield sets. For model (a) with $y/Z_\odot=1$, the two thin disk peaks are close enough together that they approximate a single peak, especially once observational uncertainties are factored in. With the \yZ{2} yield set, however, there is a $\sim0.2$ dex separation between the low- and intermediate-$\alpha$ peaks thanks to increased efficiency of CCSN element production. As a result, model (b) predicts a high density of stars at $\mathOFe\approx+0.2$ where the data show a relatively low density.

>>>>

% \todo{Hook, summary of alpha-bimodality.}
\lettrine{G}{alactic chemical evolution} (GCE) studies aim to explain the observed stellar abundance patterns in the Milky Way (MW) by modeling the star formation history and evolution of the Galaxy. A long-standing paradigm of GCE is that the metallicity of the interstellar medium (ISM) increases over time due to supernova enrichment from successive generations of stars. In this view, one feature of the MW disk that is difficult to explain is the so-called ``$\alpha$-bimodality'': the presence of two populations of stars at similar metallicity but separated by their \aFe ratio \citep[e.g.,][]{bensby_exploring_2014}. The high-$\alpha$ sequence consists of old stars \citep[$\gtrsim9$ Gyr;][]{pinsonneault_apokasc-3_2025} with super-Solar \aFe and is associated with the kinematic thick disk \citep[e.g.,][]{fuhrmann_nearby_1998}, while the low-$\alpha$ sequence is younger, with approximately Solar \aFe, and is associated with the thin disk. The $\alpha$-bimodality is present across the Galactic disk, but the relative strength of the high- and low-$\alpha$ sequences varies by location \citep{hayden_chemical_2015}.

An explanation for the MW $\alpha$-bimodality has yet to be broadly accepted in the GCE literature. An $\alpha$-bimodality is not a universal feature in simulated MW-mass galaxies \todo{(citations; also mention external galaxies?)}, so its presence and characteristics in our Galaxy may provide clues to its unique evolutionary history. GCE models that attempt to solve this problem generally fall into two camps. Some explain the $\alpha$-bimodality as a result of secular processes, such as the radial migration of stars and the inside-out growth of the disk \citep[e.g.,][]{kubryk_evolution_2015,sharma_chemical_2021,chen_chemical_2023,prantzos_origin_2023}. Others argue for a bursty star formation history, perhaps driven by multiple discrete accretion events \citep[e.g.,][]{chiappini_chemical_1997,mackereth_origin_2018,spitoni_beyond_2023} or a change in the star formation efficiency \citep[e.g.][]{conroy_birth_2022}.

% Summary of two-infall literature
The two-infall model of chemical evolution was proposed by \citet{chiappini_chemical_1997} to explain the origin of the high- and low-$\alpha$ disks. Though the model has been revised and refined since, the basic premise remains the same: the rate of gas infall onto the Galaxy is described by two consecutive, exponentially declining bursts. The relatively low infall rate between the two bursts leads to a hiatus in the star formation rate, allowing the gas abundance to evolve between the high- and low-$\alpha$ sequences while producing few stars in between. 
The infall timescale for the low-$\alpha$ disk can be varied to produce inside-out disk growth and a radial metallicity gradient \citep{romano_mass_2000}.
The initial model of \citet{chiappini_chemical_1997} successfully reproduced the available abundance data at the time, which were largely confined to the Solar neighborhood.

@@

% \todo{Outflows.} 
Most previous studies of the two-infall model have not included mass-loaded outflows. Some hydrodynamic simulations of Galactic fountains ejected by CC SNe have shown that ejected material falls back onto the disk on relatively short timescales \citep{spitoni_galactic_2008,spitoni_effects_2009} and close to their point of origin \citep{melioli_hydrodynamical_2008,melioli_hydrodynamical_2009}, suggesting the effect on GCE could be minimal. However, other simulations of MW-like galaxies with different feedback prescriptions do produce mass-loaded outflows \todo{(citations)}. Empirically, mass-loaded outflows have been observed in nearby starburst galaxies \todo{(citations)} and dwarf galaxies \todo{(citations)} but not MW-like systems, although the predicted column densities are below current detection limits \todo{(citations)}. In the absence of outflows, some two-infall studies have implemented radial gas flows or a variable star formation efficiency in order to regulate the radial metallicity gradient \citep{spitoni_effects_2011,palla_chemical_2020}.
% We adopt an {\it ad hoc} prescription for outflows as a function of radius in order to balance different yield sets and reproduce the radial metallicity gradient. \todo{Trim down, move some to methods.}

@@

% \todo{Observational evidence other than $\alpha$-bimodality.} 
The chemical and kinematic separation of the high- and low-$\alpha$ disks remains the primary observational evidence behind the two-infall model. \citet{spitoni_remind_2024} argued that the observed gap between the sequences in \aFe, contrasted with their overlap in $[\alpha/{\rm H}]$, indicates a haitus in the star formation history, which is a natural consequence of the two-infall model. In different samples, \citet{nissen_high-precision_2020} and \citet{nataf_accurate_2024} observed multiple sequences in the local age--metallicity relation, which would naturally be explained by the two-infall scenario. Many two-infall studies have also reproduced the metallicity gradient, the local surface densities of stars and gas, and the local star formation and supernova rates \todo{(citations)}, although these observables can also be matched by other models \todo{(citations)}.

% \todo{Summary of secular evolution models which claim alpha-bimodality.} 
In contrast to the two-infall scenario, a number of studies have attempted to explain the $\alpha$-bimodality through purely secular processes. 
Using a detailed prescription for radial migration, \citet{schonrich_chemical_2009} produced distinct high- and low-$\alpha$ sequences, but they did not overlap in metallicity space as would be found by later surveys \citep[e.g.,][]{bensby_exploring_2014}. Others have produced a more MW-like $\alpha$-bimdodality using a combination of radial migration and inside-out galaxy growth \citep[e.g.,][]{kubryk_evolution_2015,sharma_chemical_2021,chen_chemical_2023,prantzos_origin_2023}. In this scenario, the local high-$\alpha$ population originates from the inner Galaxy, where the star formation rate peaked early in its history. \citet{sharma_chemical_2021} and \citet{chen_chemical_2023} suggest that a simultaneous decline in the star formation rate and the \aFe ratio is needed to separate the sequences in chemical space. \citet{chen_recent_2025} additionally argue that the double sequence in the local age--metallicity relation observed by \citet{nissen_high-precision_2020} can also be explained by smooth star formation with inside-out growth. On the other hand, some GCE models that incorporate both radial migration and smooth, inside-out star formation have failed to produce an $\alpha$-bimodality \citep[e.g.,][]{johnson_stellar_2021,dubay_galactic_2024}.
% While the two-infall scenario naturally produces an $\alpha$-bimodality, it is not the only model that can do so.
While the $\alpha$-bimodality remains a key piece of evidence for the two-infall scenario, it has been reproduced by other models.

% By incorporating stellar ``churning'' (changes in guiding-center radius) and ``blurring'' (changes in orbital eccentricity) as well as radial gas flows in a multi-ring model, \citet{schonrich_chemical_2009} successfully produced distinct high- and low-$\alpha$ sequences through secular processes alone; however, they failed to reproduce the overlap in metallicity between the two that would be a key finding of later surveys \citep[e.g.,][]{bensby_exploring_2014}. 
% \citet{kubryk_evolution_2015} produce two parallel sequences in \aFe vs [Fe/H] by making a simple age cut to the results of their multi-zone GCE model, though because they do not present stellar \aFe distributions it is hard to determine whether their distributions are truly bimodal. In the model of \citet{sharma_chemical_2021}, a sharp transition in stellar \aFe for $8-12\Gyr$ old stars plus radial migration are the key ingredients for predicting a MW-like bimodality; however, their approach to the chemical evolution is not {\it ab initio}, meaning that instead of integrating the enrichment equations for their adopted star formation history, they simply assume a functional form for the abundance evolution over time \todo{(rewrite)}. \todo{Sharma basically says that it's possible to have a continuous star formation history and reproduce the $\alpha$-bimodality.} \citet{chen_chemical_2023} confirmed the findings of \citet{sharma_chemical_2021} in an {\it ab initio} multi-zone GCE model (with a two-phase ISM and radial gas flows but no outflows) and were able to produce a bimodal \aFe distribution; however, some discrepancies remain, including that the metallicity of the high-$\alpha$ sequence is too low and not extended enough. \citet{chen_chemical_2023} found that the abundance gap is caused by the simultaneous decline in SFR and \aFe. 
% \citet{prantzos_origin_2023} attribute the formation of their thick disk to a combination of inside-out disk formation and radial migration causing thick disk stars from the inner galaxy to end up in the Solar circle.
% \citet{chen_recent_2025} argue that the double age--metallicity sequences observed by \citet{nissen_high-precision_2020} can be explained by a smooth star formation history with inside-out growth: the older sequence consists of stars which have migrated from the inner galaxy, where the star formation rate peaked early, while the younger sequence was formed locally. \todo{Trim this WAY down.}

% \todo{Introduce equilibrium model.} 
As stellar age estimation techniques have improved over recent years, large catalogs have become available with ages for hundreds of thousands or even millions of stars from a wide swath of the Galaxy. In a challenge to the traditional view of GCE, which expects the ISM metallicity to continually increase over time, \citet{johnson_milky_2024} examined the age--metallicity relation at different radii from the astroNN catalog \citep{mackereth_dynamical_2019} and found that the mode of the metallicity distribution at a given radius is nearly independent from age over the past $\sim9\Gyr$. They propose an ``equilibrium scenario'' in which the local metallicity is driven by the ratio of star formation to accretion at a given radius, which evolves to a constant over $\sim\Gyr$ timescales. Whether the equilibrium metallicity is regulated by outflows, as proposed by \citet{johnson_milky_2024}, or by other factors such as a radial gas flow, the current data suggest that the gas abundance across the Galaxy has evolved very little over most of the thin disk lifetime.

% Recently, \citet{dubay_galactic_2024} studied the effect of the SN Ia delay-time distribution in the context of several different star formation histories, including a two-infall model. While the results did not greatly depend on the choice of SFH, the authors \todo{(we?)} noted that the two-infall model, as implemented, produced a stellar [O/Fe] distribution with three distinct peaks, at variance with the data.

% \todo{Tie it all together, motivate the paper.}
In light of the findings of \citet{johnson_milky_2024} and a new empirical yield scale from \citet{weinberg_scale_2024}, we evaluate the predictions of the two-infall model against stellar age and abundance data across the MW disk.
% \todo{Goal and summary of paper.} 
We run multi-zone GCE models with a two-infall accretion history, radially-dependent mass-loaded outflows, and a prescription for radial migration tuned to a hydrodynamical simulation. We investigate the impact of the scale of SN yields and outflows, the strength of radial migration, the enrichment of the circumgalactic medium, and the local disk mass surface density ratio on the GCE model predictions. We compare our results to abundance distributions across the disk from APOGEE DR17, and to age--abundance relations from multiple age catalogs. We describe our observational sample in Section \ref{sec:observational-sample}, and we detail our chemical evolution models and parameter selection in Section \ref{sec:methods}. We compare our multi-zone model predictions to the data in Section \ref{sec:multizone-results}. We discuss in Section \ref{sec:discussion} and conclude in Section \ref{sec:conclusions}.

@@

@@

The infall rate is described by two successive, exponentially declining bursts in time. The first infall component induces the formation of the thick disk, and the second component produces the thin disk. At a given galactocentric radius $R_{\rm gal}$, the un-normalized form of the infall rate is
\begin{equation}
    \label{eq:twoinfall-ifr}
    f_{\rm in}(t|R_{\rm gal}) = e^{-t/\tau_1} + f_{2/1}(R_{\rm gal}) e^{-(t-t_{\rm max})/\tau_2},
\end{equation}
where $\tau_1$ and $\tau_2$ are the first and second infall timescales, respectively, $t_{\rm max}$ is the onset of the second infall and thus the time of maximum gas infall, and $f_{2/1}$ is the ratio of the second infall amplitude to the first. We numerically calculate $f_{2/1}$ for each zone such that the resulting stellar density profile follows a two-component disk, with the surface density ratio of the thick and thin disks given by
\begin{equation}
    f_\Sigma(R) \equiv \frac{\Sigma_1(R)}{\Sigma_2(R)} = f_\Sigma(R_\odot) e^{(R-R_\odot)\cdot(1/R_2 - 1/R_1)}.
\end{equation}
We adopt a thick disk scale radius of $R_1=2.0\kpc$, a thin disk scale radius of $R_2=2.5\kpc$, and a fiducial value for the local surface density ratio of $f_\Sigma(R_\odot)=0.12$ \citep{bland-hawthorn_galaxy_2016}.

The thick-to-thin disk density ratio is especially important for our GCE models as it controls the quantity of gas accreted during each infall epoch. Our fiducial value of $f_\Sigma(R_\odot)=0.12$ is on the low end of literature estimates, which range from $f_\Sigma(R_\odot)\sim0.06-0.6$ \citep[e.g.,][]{gilmore_new_1983,siegel_star_2002,juric_milky_2008,mackereth_age-metallicity_2017,fuhrmann_local_2017}. Previous two-infall studies have adopted a similarly broad range of values (e.g., $f_\Sigma(R_\odot)=0.18$ from \citealt{spitoni_apogee_2021}; $f_\Sigma(R_\odot)=0.4$ from \citealt{spitoni_remind_2024}). We therefore explore values up to $f_\Sigma(R_\odot)=0.5$ in our multi-zone models in Section \ref{sec:multizone-results}.
% This value is on the low end of literature estimates; by comparison, \citet{mackereth_age-metallicity_2017} found $f_\Sigma(R_\odot)=0.18\pm0.05$ from APOGEE, and \citet{fuhrmann_local_2017} derived $f_\Sigma(R_\odot)=0.57$ from a very small volume (25 pc radius). 
% \citet{spitoni_galactic_2020} explores $f_\Sigma(R_\odot)\approx0.23-0.31$, \citet{palla_chemical_2020} assumes $f_\Sigma(R_\odot)=0.23$, \citet{spitoni_apogee_2021} derives $f_\Sigma(R_\odot)=0.18$ (again from APOGEE), \citet{spitoni_remind_2024} assumes $f_\Sigma(R_\odot)=0.4$.

@@

In our multi-zone models, we vary the second infall timescale with radius to produce inside-out growth of the disk. Previous multi-zone two-infall studies \citep[e.g.,][]{chiappini_abundance_2001,palla_chemical_2020} scale $\tau_2$ linearly with radius, with $\tau_2\approx1\Gyr$ in the inner disk and $\tau_2=7\Gyr$ at the Solar annulus. This prescription was adopted to match the metallicity distribution of the Solar neighborhood and the bulge in the absence of mass-loaded outflows \citep{romano_mass_2000}. We instead adopt an exponential $\tau_2 - R_{\rm gal}$ relation, with $\tau_2(R_\odot)=15\Gyr$ at the Solar annulus and a scale radius $R_{\tau_2}=7\kpc$. This is similar to the star formation history timescale of \citet{johnson_stellar_2021}, which was based on the  stellar age gradients in Milky Way-like spirals observed by \citet{sanchez_spatially_2020}. We also run models with a linear prescription and with a uniform value for $\tau_2$ and find little difference in our key results.

Finally, panel (c) of Figure \ref{fig:twoinfall-parameters} shows that the time of maximum infall $t_{\rm max}$ (c) strongly affects the overall stellar abundance distribution for values $t_{\rm max}\leq2\Gyr$, but in this case the gas tracks do not produce the characteristic abundance loop. For $t_{\rm max}>2\Gyr$, varying $t_{\rm max}$ results in a minor shift to the mean of the MDF and little change to the [O/Fe] distributions, even though the abundance tracks in [O/Fe]--[Fe/H] space appear very different. The value of $t_{\rm max}$ also slightly adjusts the ISM abundance endpoint, as a longer $t_{\rm max}$ means the chemical evolution ``reset'' from the second infall occurs closer to the present day (see discussion in Section \ref{sec:age-abundance}. We adopt a fiducial value of $t_{\rm max}=4.2\Gyr$, i.e. a lookback time of $9\Gyr$, which is generally in line with previous two-infall studies \citep[e.g.,][]{nissen_high-precision_2020,spitoni_galactic_2020,spitoni_apogee_2021}. This ensures that our models are compatible with the median age of the thick disk in the APOKASC-3 catalog of $9.14\pm0.05\Gyr$ \citep{pinsonneault_apokasc-3_2025}.

The Milky Way's last major merger with the dwarf galaxy dubbed Gaia Sausage-Enceladus \citep[GSE;][]{belokurov_co-formation_2018,helmi_merger_2018} has been proposed as an important influence on the transition from the thick disk to the thin disk, as in \citet{spitoni_remind_2024}. Our fiducial value of $t_{\rm max}=4.2\Gyr$ places the start of the formation of the thin disk close to the GSE merger (within uncertainties), which likely occurred $\sim10\Gyr$ ago \citep[e.g.,][]{helmi_merger_2018,gallart_uncovering_2019,naidu_reconstructing_2021,woody_rapid_2025}.

@@

% \begin{figure*}
%     \centering
%     \includegraphics{figures/spitoni_comparison.pdf}
%     \caption{Chemical abundance tracks and distributions from one-zone models which use our fiducial parameters (solid curves) and the MCMC-derived parameters from \citet[][dashed curves]{spitoni_apogee_2021} at three different galactocentric radii. All models use the $y/Z_\odot=1$ yield set and corresponding outflow prescription (see Table \ref{tab:yields}). The time markers and marginal panels are the same as in Figure \ref{fig:twoinfall-parameters}. We also include a 2-D heatmap of the APOGEE stellar abundances in each panel for reference. APOGEE stars are restricted to (a) $2\leq R_{\rm gal}<6$ kpc, (b) $6\leq R_{\rm gal}<10$ kpc, and (c) $10\leq R_{\rm gal}<14$ kpc, and $0\leq |z|<2$ kpc in each panel.}
%     \label{fig:spitoni-comparison}
% \end{figure*}

% \todo{Highlight differences between our models and those of \citet{spitoni_apogee_2021} in Figure \ref{fig:spitoni-comparison}.}

@@

The two-infall model generically predicts {\it three} peaks in the [O/Fe] distribution, which correspond to the high-$\alpha$ sequence, the abundance ``turn-over'' after the second infall, and finally the late-time low-$\alpha$ sequence. We previously noted this feature in \citet{dubay_galactic_2024}. Figure \ref{fig:ofe-distribution-yields} compares [O/Fe] distributions from across the Galactic disk produced by models with the \yZ{1} and \yZ{2} yield sets. For model (a) with $y/Z_\odot=1$, the two thin disk peaks are close enough together that they approximate a single peak, especially once observational uncertainties are factored in. With the \yZ{2} yield set, however, there is a $\sim0.2$ dex separation between the low- and intermediate-$\alpha$ peaks thanks to increased efficiency of CCSN element production. As a result, model (b) predicts a high density of stars at $\mathOFe\approx+0.2$ where the data show a relatively low density.
